Clustering in a hyperbolic model of complex networks.

### 2020-03-10 proofs_infinite_k.tex
Edits by Tobias

\section{Overview of the proof strategy for $k \to \infty$}

@@

We start with introducing a slightly modified version of the local clustering function, which will be convenient for computations later,
\begin{equation}\label{eq:def_local_clustering_ast_general}
	c^\ast(k;G) = \frac{1}{\Exp{N_k}} \sum_{v \in V(G) \atop \text{deg}(v)=k} c(v).
\end{equation}
Notice that the only difference between $c(k;G)$ and $c^\ast(k;G)$ is that we replace $N_k$ by its expectation $\Exp{N_k}$. The advantage is that now, the only randomness is in the formation of triangles. In addition, note that since $\Exp{N_k} > 0$ a case distinction for $N_k$ is no longer needed for $c^\ast(k;G)$. It is however still relevant since we are eventually interested in $c(k;G)$. Following the notational convention, throughout the remainder of this paper we write $c^\ast(k; \GPo)$ and $c^\ast(k; \Gbox)$ to denote the modified local clustering function in $\GPo$ and $G_{\mathcal{P},n}(\alpha,\nu)$, respectively.

@@

We then establish that the modified local clustering function for KPKVB graphs $G_n$ behaves similarly to that in the $\GPo$. The proof, found in Section~\ref{ssec:coupling_H_HP}, is based on a standard coupling between a Binomial Point Process and Poisson Point Process.

>>>>

\section{Clustering when $k\to\infty$ : overview of the proof strategy}

@@

We start with introducing a slightly modified version of the local clustering function, which will be convenient for computations later,
\begin{equation}\label{eq:def_local_clustering_ast_general}
	c^\ast(k;G) = \frac{1}{\Exp{N(k)}} \sum_{v \in V(G) \atop \text{deg}(v)=k} c(v).
\end{equation}
Notice that the only difference between $c(k;G)$ and $c^\ast(k;G)$ is that we replace $N(k)$ by its expectation $\Exp{N(k)}$. The advantage is that now, the only randomness is in the formation of triangles. In addition, note that since $\Exp{N(k)} > 0$ a case distinction for $N(k)$ is no longer needed for $c^\ast(k;G)$. It is however still relevant since we are eventually interested in $c(k;G)$. Following the notational convention, throughout the remainder of this paper we write $c^\ast(k; \GPo)$ and $c^\ast(k; \Gbox)$ to denote the modified local clustering function in $\GPo$ and $G_{\mathcal{P},n}(\alpha,\nu)$, respectively.

@@

We then establish that the modified local clustering function for KPKVB graphs $G_n$ behaves similarly to that in $\GPo$. The proof, found in Section~\ref{ssec:coupling_H_HP}, is based on a standard coupling between a Binomial Point Process and Poisson Point Process.

### 2020-03-10 transition.tex
updated section 3
Edits by Tobias

%We now turn to the task of calculating the expected number of triangles of a node at height $y$, for both the infinite model and the finite box model. 
For any $p \in \R \times \R_+$ we define for the finite box model,
\[
	T_{\text{box}}(p) = \sum^{\ne}_{p_1, p_2 \in \Pcal \setminus \{p\}} T_{\text{box}}(p,p_1,p_2)
\]
where the sum is over all distinct pairs in $\Pcal \setminus p$ and
\[
	T_{\text{box}}(p,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPon{p_1}}.
\]
Similarly, for the infinite model we define
\[
	T_\infty(y) = \sum_{p_1, p_2 \in \Pcal \setminus (0,y)}^{\ne} T_\infty(y,p_1,p_2),
\]
where
\[
	T_{\infty}(y,p_1,p_2) = \ind{p_1 \in \BallPo{y}}\ind{p_2 \in \BallPo{y}}\ind{p_2 \in \BallPo{p_1}}.
\]
Here we slightly abuse notation and write $\BallPo{y}$ for $\BallPo{(0,y)}$. We will adopt this notational convention throughout the remainder of this section, to keep notation concise. We also recall that $\Nbox(k)$ denotes the number of vertices with degree $k$ in $\Gbox$.

@@

\begin{lemma}
Let $\gamma(k_n)$ be defined as in~\eqref{eq:gammakint}. Then as $n \to \infty$
\begin{equation}\label{eq:alt_gamma_kn}
	\gamma(k_n) = (1+\smallO{1})\frac{1}{k_n^2 p_{k_n}} \int_{\Kcal_{C}(k_n)} \Exp{T_\infty(y)} \rho(y,k) 
				\alpha e^{-\alpha y} \dd y. 
\end{equation}
Moreover,
\begin{equation}\label{eq:alt_c_ast_kn}
	\Exp{c^\ast(k_n;\Gbox)} = (1 + \smallO{1}) \frac{1}{k_n^2 p_{k_n}} \int_{\Kcal_{C}(k_n)} \Exp{T_{\text{box}}(y)}
		\rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y
\end{equation}
as $n \to \infty$,
\end{lemma}

\begin{proof}
Recall that 
\[
	P(y) = \Exp{\ind{u_1 \in \BallPo{u_2}}},
\]
where $u_1$ and $u_2$ are independent and distributed according to the probability density \\$\Mu{\BallPo{y}}^{-1} \ind{u_i \in \BallPo{y}} f(x_i, y_i)$. It then follows from the Campbell-Mecke formula that
\begin{align*}
	\Exp{T_\infty(y)} &= \int \ind{p_1 \in \BallPo{y}}\ind{p_2 \in \BallPo{y}}
		\ind{p_2 \in \BallPo{p_1}} f(x_1,y_1) f(x_2,y_2) \dd x_1 \dd x_2 \dd y_1 \dd y_2\\
	&= \Mu{\BallPo{y}}^2 P(y).
\end{align*}
It then follows that,
\begin{align*}
	\gamma(k_n) &= \frac{1}{p_{k_n}} \cdot \int_0^\infty P(y) \rho(y,k) \alpha e^{-\alpha y} \dd y
	= \frac{1}{p_{k_n}} \int_0^\infty \Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} \rho(y,k) 
		\alpha e^{-\alpha y} \dd y.
\end{align*}

Because $\Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} = \bigO{1}$, uniformly in $y \in \R_+$ and $\mu(y) = (1+\smallO{1}) k_n$ uniformly for $y \in \Kcal_C(k_n)$, by the concentration of heights (Proposition~\ref{prop:concentration_height_general})
\begin{align*}
	&\hspace{-30pt} \frac{1}{p_{k_n}} 
		\int_0^\infty \Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} \rho(y,k) \alpha e^{-\alpha y} \dd y \\
	&= (1+\smallO{1}) \frac{1}{p_{k_n}} 
		\int_{\Kcal_C(k_n)} \Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} \rho(y,k) \alpha e^{-\alpha y} \dd y \\
	&= (1+ \smallO{1}) \frac{1}{k_n^2 p_{k_n}} \int_{\Kcal_C(k_n)} \Exp{T_\infty(y)} \rho(y,k) \alpha e^{-\alpha y} \dd y.
\end{align*}

For~\eqref{eq:alt_c_ast_kn} we recall that
\[
	c^\ast(k_n; \Gbox) = \frac{1}{\Exp{\Nbox(k_n)}} \sum_{p \in \Pcal} c_{\text{box}}(p)\ind{\Dbox(p) = k_n},
\]
where $c_{\text{box}}(p)$ can be expressed as
\[
	c_{\text{box}}(p) = \frac{1}{\binom{\Dbox(p)}{2}} \sum_{p_1, p_2 \in \Pcal \setminus p}^{\ne} T_{\text{box}}(p,p_1,p_2)
	= \frac{T_{\text{box}}(p)}{\binom{\Dbox(p)}{2}}.
\]
By the Campbell-Mecje formula
\begin{align*}
	\Exp{c^\ast(k_n; \Gbox)} 
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \Exp{c_{\text{box}}(p)\ind{\Dbox(p) = k_n}} f(x,y) \dd x \dd y\\
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \CExp{c_{\text{box}}(p)}{\Dbox(p) = k_n}\rho_{\text{box}}(p,k_n) 
		f(x,y) \dd x \dd y\\
	&= (1+\smallO{1})\frac{n}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
		\rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{align*}
where the last line follows from the concentration of heights, for which we used the upper bound $\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n} \le 1$.

To analyze the conditional expectation we observe that, similar to the analysis of $\gamma(k_n)$, conditioned on there being $k_n$ points in $\BallPon{y}$, each point $u_i = (x_i,y_i)$ is independently distributed according to the probability density $\Mu{\BallPon{y}}^{-1} \ind{u_i \in \BallPon{y}} f(x_i,y_i)$. Therefore,
\begin{align*}
	\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
	&= \binom{k_n}{2}^{-1} \Exp{\sum_{1 \le i < j \le k_n} \ind{u_i \in \BallPon{u_j}}} \\
	&= \Exp{u_1 \in \BallPon{u_2}} \\
	&= \Mu{\BallPon{y}}^{-2} \iint T_{\text{box}}(y,p_1,p_2) f(x_1,y_1) f(x_2,y_2) 
		\dd x_1 \dd y_1 \dd x_2 \dd y_2\\
	&= \Mu{\BallPon{y}}^{-2} \Exp{T_{\text{box}}(y)}.
\end{align*}
and thus, by applying a concentration of heights argument on $\Mu{\BallPon{y}}^{-2}$,
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{n \Mu{\BallPon{2\log(k_n/\xi)}}^{-2}}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \Exp{T_{\text{box}}(y)} \rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
To finish the argument, we first note that $\Mu{\BallPon{2\log(k_n/\xi)}}^{-2} = (1+\smallO{1})k_n^2$, while
\[
	\Exp{N_{\text{box}}(k_n)} = (1+\smallO{1}) n p_{k_n}
\]
by Lemma~\ref{lem:average_degree_G_box}. We therefore conclude that
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{1}{k_n^2 p_{k_n}} \int_{\Kcal_C(k_n)} 
		\Exp{T_{\text{box}}(y)} \rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
\end{proof}

@@

%Comparing~\eqref{eq:alt_gamma_kn} and~\eqref{eq:alt_c_ast_kn}, we conclude that to prove Proposition~\ref{prop:convergence_average_clustering_P_n} it is enough to show that
%\begin{equation}\label{eq:triangle_diff_main_term}
%	 \left|\int_{\Kcal_{C}(k_n)} \Exp{T_{\text{box}}(y) - T_\infty(y)}
%	\rho(y,k) \alpha e^{-\alpha y} \, d y \right|
%	= \smallO{s(k_n)p_{k_n}k_n^2},
%\end{equation}
%which means we have to compute the expected difference in triangles between both models.

@@

Let us now define the following triangle count function
\[
	\widetilde{T}_{\text{box}}(p_0) = \sum_{(p_1, p_2) \in \Pcal \setminus \{p_0\}}^{\ne} 
		\widetilde{T}_{\text{box}}(p_0,p_1,p_2).
\]
where
\[
	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal}.
\]
Then $\widetilde{T}_{\text{box}}(p_0)$ only counts those triangles attached to $p_0$ that exist in both $\Gbox$ and $\Ginf$ and thus, by definition of the region $\mathcal{T}(p_0,p_1)$,
\[
	T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)
	= \sum_{p_1, p_2 \in \Pcal \setminus \{p_0\}}^{\ne} \ind{p_1 \in \BallPon{p_0}} 
			\ind{p_2 \in \mathcal{T}(p_0, p_1)}.
\]
%Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.
%Observe that the difference between $\widetilde{T}_{\text{box}}(p_0)$ and $T_{\text{box}}(p_0)$ is in the last indicator in $\widetilde{T}_{\text{box}}(p_0,p_1,p_2)$.  counts those points $p_2$ in the box $\Rcal$ that are in the neighbourhood $\BallPo{p_1}$ of $p_1$ in the infinite model. Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.

@@

Define $\Rcal^\prime := (\R \times \R_+) \setminus \Rcal$. Then we can write
\begin{align*}
	\left|T_{\mathrm{box}}(p_0) - T_\infty(y)\right|
	&\le \left|T_{\mathrm{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right| 
		+ \left|\widetilde{T}_{\text{box}}(p_0) - T_\infty(y)\right|\\
	&= \left|T_{\mathrm{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right| 
		+ \sum_{p_1, p_2 \in \Pcal \cap \Rcal^\prime}^{\ne} T_{\infty}(p_0,p_1,p_2).
\end{align*}
Then by the Campbell-Mecke formula
\begin{align*}
	\left|\Exp{T_{\text{box}}(p_0) - T_{\infty}(p_0)}\right|
	&\le \Exp{\left|T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right|}\\
	&\hspace{10pt}+ \int_{\Rcal^\prime}\int_{\Rcal^\prime} T_{\infty}(p_0,p_1,p_2) f(x_1,y_1) f(x_2,y_2)
		\dd x_2 \dd y_2 \dd x_1 \dd y_1. 
\end{align*}

@@

	\Exp{c^\ast(k_n;\Gbox)} &= (1 + \smallO{1}) \frac{1}{p_{k_n}} \int_{\Kcal_{C}(k_n)} P(y)

@@

	&\hspace{10pt}+ \smallO{s(k_n)}	\frac{1}{p_{k_n}} \int_{\Kcal_{C}(k_n)} P(y)

@@

By Lemma~\ref{lem:degree_integral} the integral in the second term is $(1+\smallO{1}) p_{k_n}$ and thus the second term is $\smallO{s(k_n)} = \smallO{\gamma(k_n)}$. Hence it remain to prove that the first term is $(1+\smallO{1}) \gamma(y)$. Using~\eqref{eq:alt_gamma_kn} it is enough to show that

@@

%	&= \bigO{1} \left(\log(k_n) n^{-(2\alpha - 1)} + k_n^2 n^{-(4\alpha - 2)} \right) p_{k_n} = \smallO{s(k_n) p_{k_n} k_n^2},

@@

%\begin{proof}
%We first write
%\[
%	\Exp{\left|\widetilde{T}_{\text{box}}(y) - T_{\infty}(y)\right|} 
%	\le \Exp{\left|T_{\text{box}}(y) - \widetilde{T}_{\text{box}}(y) \right|}
%	+\Exp{\left|T_{\text{box}}(y) - T_{\infty}(y)\right|}.
%\]
%Therefore, Lemma~\ref{lem:clustering_error_T_term} and equation~\eqref{eq:triangle_count_diff_scaling} imply that, uniformly for $y \in \Kcal_C(k_n)$,
%\[
%	\Exp{\left|\widetilde{T}_{\text{box}}(y) - T_{\infty}(y)\right|} = \bigO{\log(k_n)n^{-(2\alpha - 1)}
%	+ k_n^2 n^{-(4\alpha - 2)}}	= \smallO{s(k_n) k_n^2},
%\]
%where the last part is due to Lemma~\ref{lem:scaling_s_alpha}.
%Next, since $\Exp{T_\infty(y)} = \Mu{\BallPo{y}}^2 P(y)$, we get
%\[
%	\Exp{\widetilde{T}_{\text{box}}(y)} = \Exp{T_{\infty}(y)} + \Exp{\widetilde{T}_{\text{box}}(y) - T_{\infty}(y)} 
%	= k_n^2 P(y) + \smallO{s(k_n)k_n^2},
%\]
%uniformly on $\Kcal_{C}(k_n)$. Therefore, we can apply a concentration of height argument to replace $\rho_{\text{box}}(y,k_n)$ with $\rho(y,k_n)$ and thus obtain
%\begin{align*}
%	&\hspace{-30pt}\int_{\Kcal_{C}(k_n)} \rho_{\text{box}}(y,k_n) \Exp{\widetilde{T}_{\text{box}}(y)} 
%		f(x,y) \dd x \dd y\\
%	&= n k_n^2 \int_{\Kcal_{C}(k_n)} \rho(y,k_n) \left(P(y) + \smallO{s(k_n)}\right)
%		\alpha e^{-\alpha y} \dd y\\
%	&= (1+\smallO{1}) n k_n^2 \int_{\Kcal_{C}(k_n)} P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y\\
%	&= (1+\smallO{1}) n k_n^2 \int_0^\infty P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
%\end{align*}
%where we used that on $\Kcal_{C}(k_n)$, $P(y) = \bigT{s(k_n)}$. This proves the first statement. The second statement follows by observing that
%\[
%	\int_0^\infty P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y = p_{k_n} \gamma(k_n)
%	= \bigT{k_n^{-(2\alpha + 1)} s(k_n)}.
%\]
%\end{proof}

@@

>>>>

%We now turn to the task of calculating the expected number of triangles of a node at height $y$, for both the infinite model and the finite box model. 
For any $p \in \R \times \R_+$ we define for the finite box model,
\[
	T_{\text{box}}(p) = \sum_{p_1, p_2 \in \Pcal \setminus \{p\}, \atop \text{distinct}} T_{\text{box}}(p,p_1,p_2)
\]
where the sum is over all distinct pairs in $\Pcal \setminus p$ and
\[
	T_{\text{box}}(p,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPon{p_1}}.
\]
Similarly, for the infinite model we define
\[
	T_\infty(y) = \sum_{p_1, p_2 \in \Pcal \setminus (0,y), \atop \text{distinct}} T_\infty(y,p_1,p_2),
\]
where
\[
	T_{\infty}(y,p_1,p_2) = \ind{p_1 \in \BallPo{y}}\ind{p_2 \in \BallPo{y}}\ind{p_2 \in \BallPo{p_1}}.
\]
Here we slightly abuse notation and write $\BallPo{y}$ for $\BallPo{(0,y)}$. We will adopt this notational convention throughout the remainder of this section, to keep notation concise. We also recall that $\Nbox(k)$ denotes the number of vertices with degree $k$ in $\Gbox$.

@@

\begin{lemma}
Let $\gamma(k_n)$ be defined as in~\eqref{eq:gammakint}. Then as $n \to \infty$
\begin{equation}\label{eq:alt_gamma_kn}
	\gamma(k_n) = (1+\smallO{1})\frac{1}{k_n^2 \pmf(k_n)} \int_{\Kcal_{C}(k_n)} \Exp{T_\infty(y)} \rho(y,k) 
				\alpha e^{-\alpha y} \dd y. 
\end{equation}
Moreover,
\begin{equation}\label{eq:alt_c_ast_kn}
	\Exp{c^\ast(k_n;\Gbox)} = (1 + \smallO{1}) \frac{1}{k_n^2 \pmf(k_n)} \int_{\Kcal_{C}(k_n)} \Exp{T_{\text{box}}(y)}
		\rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y
\end{equation}
as $n \to \infty$,
\end{lemma}

\begin{proof}
Recall that 
\[
	P(y) = \Exp{\ind{u_1 \in \BallPo{u_2}}},
\]
where $u_1$ and $u_2$ are independent and distributed according to the probability density \\$\Mu{\BallPo{y}}^{-1} \ind{u_i \in \BallPo{y}} f(x_i, y_i)$. It then follows from the Campbell-Mecke formula that
\begin{align*}
	\Exp{T_\infty(y)} &= \int \ind{p_1 \in \BallPo{y}}\ind{p_2 \in \BallPo{y}}
		\ind{p_2 \in \BallPo{p_1}} f(x_1,y_1) f(x_2,y_2) \dd x_1 \dd x_2 \dd y_1 \dd y_2\\
	&= \Mu{\BallPo{y}}^2 P(y).
\end{align*}
It then follows that,
\begin{align*}
	\gamma(k_n) &= \frac{1}{\pmf(k_n)} \cdot \int_0^\infty P(y) \rho(y,k) \alpha e^{-\alpha y} \dd y
	= \frac{1}{\pmf(k_n)} \int_0^\infty \Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} \rho(y,k) 
		\alpha e^{-\alpha y} \dd y.
\end{align*}

Because $\Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} = \bigO{1}$, uniformly in $y \in \R_+$ and $\mu(y) = (1+\smallO{1}) k_n$ uniformly for $y \in \Kcal_C(k_n)$, by the concentration of heights (Proposition~\ref{prop:concentration_height_general})
\begin{align*}
	&\hspace{-30pt} \frac{1}{\pmf(k_n)} 
		\int_0^\infty \Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} \rho(y,k) \alpha e^{-\alpha y} \dd y \\
	&= (1+\smallO{1}) \frac{1}{\pmf(k_n)} 
		\int_{\Kcal_C(k_n)} \Exp{T_\infty(y)} \Mu{\BallPo{y}}^{-2} \rho(y,k) \alpha e^{-\alpha y} \dd y \\
	&= (1+ \smallO{1}) \frac{1}{k_n^2 \pmf(k_n)} \int_{\Kcal_C(k_n)} \Exp{T_\infty(y)} \rho(y,k) \alpha e^{-\alpha y} \dd y.
\end{align*}

For~\eqref{eq:alt_c_ast_kn} we recall that
\[
	c^\ast(k_n; \Gbox) = \frac{1}{\Exp{\Nbox(k_n)}} \sum_{p \in \Pcal} c_{\text{box}}(p)\ind{\Dbox(p) = k_n},
\]
where $c_{\text{box}}(p)$ can be expressed as
\[
	c_{\text{box}}(p) = \frac{1}{\binom{\Dbox(p)}{2}} \sum_{p_1, p_2 \in \Pcal \setminus p, \atop \text{distinct}} T_{\text{box}}(p,p_1,p_2)
	= \frac{T_{\text{box}}(p)}{\binom{\Dbox(p)}{2}}.
\]
By the Campbell-Mecke formula
\begin{align*}
	\Exp{c^\ast(k_n; \Gbox)} 
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \Exp{c_{\text{box}}(p)\ind{\Dbox(p) = k_n}} f(x,y) \dd x \dd y\\
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \CExp{c_{\text{box}}(p)}{\Dbox(p) = k_n}\rho_{\text{box}}(p,k_n) 
		f(x,y) \dd x \dd y\\
	&= (1+\smallO{1})\frac{n}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
		\rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{align*}
where the last line follows from the concentration of heights, for which we used the upper bound $\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n} \le 1$.

To analyze the conditional expectation we observe that, similar to the analysis of $\gamma(k_n)$, conditioned on there being $k_n$ points in $\BallPon{y}$, each point $u_i = (x_i,y_i)$ is independently distributed according to the probability density $\Mu{\BallPon{y}}^{-1} \ind{u_i \in \BallPon{y}} f(x_i,y_i)$. Therefore,
\begin{align*}
	\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
	&= \binom{k_n}{2}^{-1} \Exp{\sum_{1 \le i < j \le k_n} \ind{u_i \in \BallPon{u_j}}} \\
	&= \Exp{u_1 \in \BallPon{u_2}} \\
	&= \Mu{\BallPon{y}}^{-2} \iint T_{\text{box}}(y,p_1,p_2) f(x_1,y_1) f(x_2,y_2) 
		\dd x_1 \dd y_1 \dd x_2 \dd y_2\\
	&= \Mu{\BallPon{y}}^{-2} \Exp{T_{\text{box}}(y)}.
\end{align*}
and thus, by applying a concentration of heights argument on $\Mu{\BallPon{y}}^{-2}$,
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{n \Mu{\BallPon{2\log(k_n/\xi)}}^{-2}}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \Exp{T_{\text{box}}(y)} \rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
To finish the argument, we first note that $\Mu{\BallPon{2\log(k_n/\xi)}}^{-2} = (1+\smallO{1})k_n^2$, while
\[
	\Exp{N_{\text{box}}(k_n)} = (1+\smallO{1}) n \pmf(k_n)
\]
by Lemma~\ref{lem:average_degree_G_box}. We therefore conclude that
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{1}{k_n^2 \pmf(k_n)} \int_{\Kcal_C(k_n)} 
		\Exp{T_{\text{box}}(y)} \rho_{\mathrm{box}}(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
\end{proof}

@@

%Comparing~\eqref{eq:alt_gamma_kn} and~\eqref{eq:alt_c_ast_kn}, we conclude that to prove Proposition~\ref{prop:convergence_average_clustering_P_n} it is enough to show that
%\begin{equation}\label{eq:triangle_diff_main_term}
%	 \left|\int_{\Kcal_{C}(k_n)} \Exp{T_{\text{box}}(y) - T_\infty(y)}
%	\rho(y,k) \alpha e^{-\alpha y} \, d y \right|
%	= \smallO{s(k_n)\pmf(k_n)k_n^2},
%\end{equation}
%which means we have to compute the expected difference in triangles between both models.

@@

Let us now define the following triangle count function
\[
	\widetilde{T}_{\text{box}}(p_0) = \sum_{(p_1, p_2) \in \Pcal \setminus \{p_0\}, \atop \text{distinct}}
		\widetilde{T}_{\text{box}}(p_0,p_1,p_2).
\]
where
\[
	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal}.
\]
Then $\widetilde{T}_{\text{box}}(p_0)$ only counts those triangles attached to $p_0$ that exist in both $\Gbox$ and $\Ginf$ and thus, by definition of the region $\mathcal{T}(p_0,p_1)$,
\[
	T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)
	= \sum_{p_1, p_2 \in \Pcal \setminus \{p_0\}, \atop \text{distinct}} \ind{p_1 \in \BallPon{p_0}} 
			\ind{p_2 \in \mathcal{T}(p_0, p_1)}.
\]
%Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.
%Observe that the difference between $\widetilde{T}_{\text{box}}(p_0)$ and $T_{\text{box}}(p_0)$ is in the last indicator in $\widetilde{T}_{\text{box}}(p_0,p_1,p_2)$.  counts those points $p_2$ in the box $\Rcal$ that are in the neighbourhood $\BallPo{p_1}$ of $p_1$ in the infinite model. Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.

@@

Define $\Rcal^\prime := (\R \times \R_+) \setminus \Rcal$. Then we can write
\begin{align*}
	\left|T_{\mathrm{box}}(p_0) - T_\infty(y)\right|
	&\le \left|T_{\mathrm{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right| 
		+ \left|\widetilde{T}_{\text{box}}(p_0) - T_\infty(y)\right|\\
	&= \left|T_{\mathrm{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right| 
		+ \sum_{p_1, p_2 \in \Pcal \cap \Rcal^\prime, \atop \text{distinct}} T_{\infty}(p_0,p_1,p_2).
\end{align*}
Then by the Campbell-Mecke formula
\begin{align*}
	\left|\Exp{T_{\text{box}}(p_0) - T_{\infty}(p_0)}\right|
	&\le \Exp{\left|T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right|}\\
	&\hspace{10pt}+ \int_{\Rcal^\prime}\int_{\Rcal^\prime} T_{\infty}(p_0,p_1,p_2) f(x_1,y_1) f(x_2,y_2)
		\dd x_2 \dd y_2 \dd x_1 \dd y_1. 
\end{align*}

@@

	\Exp{c^\ast(k_n;\Gbox)} &= (1 + \smallO{1}) \frac{1}{\pmf(k_n)} \int_{\Kcal_{C}(k_n)} P(y)

@@

	&\hspace{10pt}+ \smallO{s(k_n)}	\frac{1}{\pmf(k_n)} \int_{\Kcal_{C}(k_n)} P(y)

@@

By Lemma~\ref{lem:degree_integral} the integral in the second term is $(1+\smallO{1}) \pmf(k_n)$ and thus the second term is $\smallO{s(k_n)} = \smallO{\gamma(k_n)}$. Hence it remain to prove that the first term is $(1+\smallO{1}) \gamma(y)$. Using~\eqref{eq:alt_gamma_kn} it is enough to show that

@@

%	&= \bigO{1} \left(\log(k_n) n^{-(2\alpha - 1)} + k_n^2 n^{-(4\alpha - 2)} \right) \pmf(k_n) = \smallO{s(k_n) \pmf(k_n) k_n^2},

@@

%\begin{proof}
%We first write
%\[
%	\Exp{\left|\widetilde{T}_{\text{box}}(y) - T_{\infty}(y)\right|} 
%	\le \Exp{\left|T_{\text{box}}(y) - \widetilde{T}_{\text{box}}(y) \right|}
%	+\Exp{\left|T_{\text{box}}(y) - T_{\infty}(y)\right|}.
%\]
%Therefore, Lemma~\ref{lem:clustering_error_T_term} and equation~\eqref{eq:triangle_count_diff_scaling} imply that, uniformly for $y \in \Kcal_C(k_n)$,
%\[
%	\Exp{\left|\widetilde{T}_{\text{box}}(y) - T_{\infty}(y)\right|} = \bigO{\log(k_n)n^{-(2\alpha - 1)}
%	+ k_n^2 n^{-(4\alpha - 2)}}	= \smallO{s(k_n) k_n^2},
%\]
%where the last part is due to Lemma~\ref{lem:scaling_s_alpha}.
%Next, since $\Exp{T_\infty(y)} = \Mu{\BallPo{y}}^2 P(y)$, we get
%\[
%	\Exp{\widetilde{T}_{\text{box}}(y)} = \Exp{T_{\infty}(y)} + \Exp{\widetilde{T}_{\text{box}}(y) - T_{\infty}(y)} 
%	= k_n^2 P(y) + \smallO{s(k_n)k_n^2},
%\]
%uniformly on $\Kcal_{C}(k_n)$. Therefore, we can apply a concentration of height argument to replace $\rho_{\text{box}}(y,k_n)$ with $\rho(y,k_n)$ and thus obtain
%\begin{align*}
%	&\hspace{-30pt}\int_{\Kcal_{C}(k_n)} \rho_{\text{box}}(y,k_n) \Exp{\widetilde{T}_{\text{box}}(y)} 
%		f(x,y) \dd x \dd y\\
%	&= n k_n^2 \int_{\Kcal_{C}(k_n)} \rho(y,k_n) \left(P(y) + \smallO{s(k_n)}\right)
%		\alpha e^{-\alpha y} \dd y\\
%	&= (1+\smallO{1}) n k_n^2 \int_{\Kcal_{C}(k_n)} P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y\\
%	&= (1+\smallO{1}) n k_n^2 \int_0^\infty P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
%\end{align*}
%where we used that on $\Kcal_{C}(k_n)$, $P(y) = \bigT{s(k_n)}$. This proves the first statement. The second statement follows by observing that
%\[
%	\int_0^\infty P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y = \pmf(k_n) \gamma(k_n)
%	= \bigT{k_n^{-(2\alpha + 1)} s(k_n)}.
%\]
%\end{proof}

@@

%Note that $\hat{y}_{\text{left}}(p,p^\prime)$ and $\hat{x}_{\text{left}}(p,p^\prime)$ correspond to, respectively, $\hat{y}(p,p^\prime)$ and $\hat{x}(p,p^\prime)$ considered in Section~\ref{ssec:missing_triangles}. The derivation of $\hat{y}_{\text{right}}(p,p^\prime)$ and $\hat{x}_{\text{right}}(p,p^\prime)$ is done in a similar manner and we omit the details here.
